Studying public transit API query logs\\ to get an indication of travel flows

### 2016-01-07 paper.tex

\rv{Starting here, this section isn't really going anywhere, at least not to a point. Consider removing the next two paragraphs and start talking about your data.}
OD type representations are used when the origin and destination preside over intermediate locations on the path of movement. 
The OD Matrix is most common visualization technique in this type of representation. 
Rows and columns correspond to locations while cells are coloured to express the volume of flow.
OD matrices can be reordered to emphasise connectivity between locations.
While this feature is particularly useful for identifying trends and outliers in the data, the tabular display is challenging for lay users to interpret and less expressive of connections between locations than representations based on the node-link metaphor.
In recent years, such representations have become increasingly prominent for  displaying movement in scientific publications as well as popular culture.
Locations are depicted as nodes and arrows are drawn between them to indicate movement in a certain direction, allowing for exact and individual relationships to be traced.

@@

Matching pairs of origin and destination (OD) stations must be aggregated prior to visualization.
Several stages of data processing are required to arrive at this outcome.
As we were specifically interested in studying morning and evening rush hour travel that occur on weekdays, we filtered the dataset to exclude data records created outside the time range of 06:00 hrs and 10:00 hrs as well as 17:00 hrs and 21:00 hrs on weekdays, and all data records created on weekend.
\rv{why? and is it just you, or also the reader?}
\todo{Alvin: no longer a problem since i changed the text to remove any mention of exact hours. PC: we will have to explain ourselves why we'd choose different hours than what we've found further on: could you change the intervals int he visualizations to what matches \cref{fig:average}?} 
While queries made by automated user agents like search engine bots and data harvesters can be considered as valid traffic, they may may over represent connections between certain stations. In this regard, such queries have been excluded to improve the accuracy of our analysis. 
\rv{Be a bit more precise here: they are valid traffic, but not helpful for prediction of human behavior.}
Similarly, data records with unparsable origin or destination station names due to invalid character encoding were removed. 
One distinctive challenge is the substantial amount of inconsistency in the way station names are logged. 
Spelling mistakes, and the use of unstandardized abbreviations in particular, reduce the quality and accuracy of the resulting visualization.
To overcome this challenge, an open-source reconciliator (\url{https://github.com/irail/stations}) in PHP was employed to reduce spelling variations and link this with a geographical location.

@@

There are several inconsistent patterns in the data that do not correspond with an existing understanding of flows in Belgium.
In particular, the large number of trips made between Antwerp city and other stations within the province of Antwerp in the morning, is not observable in other major cities in Flanders.
Further investigation led to the identification of an isolated connection originating from Antwerp city towards a suburb, suggesting anomalous usage of the iRail API that were not flagged as bots or a high concentration people who utilise route planning applications on a regular basis.
\rv{Earlier, you said you filtered bots. Is this still the case?}
Accordingly, the reverse is identified in Liege where an unusually large connection originating from a suburb towards Liege city prompted speculation as well as discussions on how such outliers should be treated or what may have led to their existence.

>>>>

OD type representations are used when the origin and destination preside over intermediate locations on the path of movement. 
The OD Matrix is most common visualization technique in this type of representation. 
Rows and columns correspond to locations while cells are coloured to express the volume of flow.
OD matrices can be reordered to emphasise connectivity between locations.
While this feature is particularly useful for identifying trends and outliers in the data, the tabular display is challenging for lay users to interpret and less expressive of connections between locations than representations based on the node-link metaphor.
In recent years, such representations have become increasingly prominent for  displaying movement in scientific publications as well as popular culture.
Locations are depicted as nodes and arrows are drawn between them to indicate movement in a certain direction, allowing for exact and individual relationships to be traced.

@@

Matching pairs of origin and destination (OD) stations must be aggregated prior to visualization.
Several stages of data processing are required to arrive at this outcome.
As we were specifically interested in studying morning and evening rush hour travel that occur on weekdays, we filtered the dataset to exclude data records created outside the time range of 06:00 hrs and 10:00 hrs as well as 17:00 hrs and 21:00 hrs on weekdays, and all data records created on weekend.
While queries made by automated user agents like search engine bots and data harvesters can be considered as valid traffic, they may may over represent connections between certain stations. In this regard, such queries have been excluded to improve the accuracy of our analysis. 
Similarly, data records with unparsable origin or destination station names due to invalid character encoding were removed. 
One distinctive challenge is the substantial amount of inconsistency in the way station names are logged. 
Spelling mistakes, and the use of unstandardized abbreviations in particular, reduce the quality and accuracy of the resulting visualization.
To overcome this challenge, an open-source reconciliator (\url{https://github.com/irail/stations}) in PHP was employed to reduce spelling variations and link this with a geographical location.

@@

There are several inconsistent patterns in the data that do not correspond with an existing understanding of flows in Belgium.
In particular, the large number of trips made between Antwerp city and other stations within the province of Antwerp in the morning, is not observable in other major cities in Flanders.
Further investigation led to the identification of an isolated connection originating from Antwerp city towards a suburb, suggesting anomalous usage of the iRail API that were not flagged as bots. Alternatively the large number of trips may have originated from users who repeatedly request for updates on delayed trains or from a concentration of people who utilise route planning applications on a regular basis.
Accordingly, the reverse is identified in Liege where an unusually large connection originating from a suburb towards Liege city prompted speculation as well as discussions on how such outliers should be treated or what may have led to their existence.
\todo{not sure if we should move this somewhere else}
As with any form of analysis based on appropriation of data, we acknowledge that more work is required to identify caveats and understand the validity of the findings yielded.
